---
title: "37810 Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##section 4.1.1
###exercises 2(a)

```{r}
a=1
a[2]=1
for (i in 3:30){
  a[i]=a[i-1]+a[i-2]
}
b=1
b[2:29]=a[3:30]/a[2:29]
print(b)
print("The sequence appear to be converging")
```

###exercises 2(b)

```{r}
c=1
c[1:29]=b[1:29]-(1+5^0.5)/2
print(c)
print("The sequence converging to golden ratio")
```

###exercises 3(a)

answer=15

```{r}
answer <- 0
for (j in 1:5) answer <- answer + j
print(answer)
```

###exercises 3(b)

answer=c(1,2,3,4,5)

```{r}
answer <- NULL
for (j in 1:5) answer <- c(answer, j)
print(answer)
```

###exercises 3(c)

answer=c(0,1,2,3,4,5)

```{r}
answer <- 0
for (j in 1:5) answer <- c(answer, j)
print(answer)
```

###exercises 3(d)

answer=120

```{r}
answer <- 1
for (j in 1:5) answer <- answer * j
print(answer)
```

###exercises 3(e)

answer=c(3,21,23,6,11,15,12,22,30,24,13,29,17,26,27,3)

```{r}
answer <- 3
for (j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
print(answer)
```

##section 4.1.2

###exercises 4

```{r}
interest=function(p,n){
  if (n<=3){
    I=p*((1+0.04)^n-1)
  }else{
    I=p*((1+0.05)^n-1)
  }
  return(I)
}
```

###exercises 5

```{r}
payment=function(n,p,open){
  if(open==TRUE){
    R=p*0.005/(1-(1+0.005)^(-n))
  }else{
    R=p*0.004/(1-(1+0.004)^(-n))
  }
  return(R)
}
```

##section 4.1.3

###exercises 2

```{r}
Fibonacci=c(1,1)
while(Fibonacci[length(Fibonacci)]+Fibonacci[length(Fibonacci)-1]<300){
  Fibonacci=c(Fibonacci,Fibonacci[length(Fibonacci)]+Fibonacci[length(Fibonacci)-1])
}
print(Fibonacci)
```

###exercises 4

```{r}
rate=function(i){
  j=0
  while(abs(i-j)>=0.000001){
    j=i
    i=(1-(1+i)^(-20))/19
  }
  return(i)
}
rate(0.006)
for(i in 1:20){
  print(rate(i*0.001))
}
```

The final interest rates would be low if the starting guesses is between 0.002 and 0.005. Otherwise, the final interest rates would be somewhat higher.

###exercises 5

```{r}
times=function(i){
  n=0
  j=0
  while(abs(i-j)>=0.000001){
    n=n+1
    j=i
    i=(1-(1+i)^(-20))/19
  }
  return(n)
}
times(0.006)

```

##section 4.1.5

###exercise 2(a)

```{r}
Eratosthenes <- function(n) {
# Print prime numbers up to n (based on the sieve of Eratosthenes)
  if (n >= 2) {
  sieve <- seq(2, n)
  primes <- c()
  while (length(sieve) > 0) {
  p <- sieve[1]
  primes <- c(primes, p)
  sieve <- sieve[(sieve %% p) != 0]
  }
return(primes)
  } else {
stop("Input value of n should be at least 2.")
  }
}
Eratosthenes(30)
```

###exercise 2(b)

```{r}
Eratosthenes <- function(n) {
# Print prime numbers up to n (based on the sieve of Eratosthenes)
  if (n >= 2) {
  sieve <- seq(2, n)
  primes <- c()
  while (length(sieve) > 0) {
  p <- sieve[1]
  if(p >= sqrt(n)){
    return(sieve)
  }
  primes <- c(primes, p)
  sieve <- sieve[(sieve %% p) != 0]
  }
  } else {
stop("Input value of n should be at least 2.")
  }
}
Eratosthenes(30)
```

###exercise 2(c)

```{r}
Eratosthenes <- function(n) {
# Print prime numbers up to n (based on the sieve of Eratosthenes)
  if (n >= 2) {
  sieve <- seq(2, n)
  primes <- c()
  while (length(sieve) > 0) {
  p <- sieve[1]
  if(p >= sqrt(n)){
    break
  }
  primes <- c(primes, p)
  sieve <- sieve[(sieve %% p) != 0]
  }
  return(c(primes,sieve))
  } else {
stop("Input value of n should be at least 2.")
  }
}
Eratosthenes(30)
```

##section 4.2.1

###exercises 2(a)

```{r}
compound.interest=function(P,i.r,n){
  return(P*(1+i.r)^n)
}
```

###exercises 2(b)

```{r}
print(compound.interest(1000,0.01,30))
```

###exercises 3

```{r}
f=function(x)x^2-1
zero=function(a,b,eps=1e-6){
  if (f(a)*f(b)>0){
    stop("Initial values should have opposite signs")
  }else{
    while(b-a>eps){
      x=(a+b)/2
      if(f(x)==0)break
      if(f(a)*f(x)<0) b=x
      else a=x
    }
  }
  return(x)
}
zero(0,3,1e-6)
```


##section 4.4.1

###exercises 1

```{r}
mergesort <- function (x,decreasing=FALSE) {
  len <-length(x)
  if (len < 2){
    result <- x
  }else {
    y <- x[1:floor(len / 2)]
    z <- x[floor(len / 2 + 1):len]
    if(decreasing==FALSE){
      y <- mergesort(y)
      z <- mergesort(z)
    }else{
      y <- mergesort(y,TRUE)
      z <- mergesort(z,TRUE)
    }
    result <- c()
    while (min(length(y), length(z)) > 0) {
      if (y[1] < z[1] && decreasing==FALSE) {
        result <- c(result, y[1])
        y <- y[-1]
      } else if(y[1] >= z[1] && decreasing==FALSE) {
        result <- c(result, z[1])
        z <- z[-1]
      } else if(y[1] > z[1] && decreasing==TRUE) {
        result <- c(result,y[1])
        y <- y[-1]
      } else {
        result <- c(result, z[1])
        z <- z[-1]
      }
      
    }
    if (length(y) > 0)
    result <- c(result, y)
    else
    result <- c(result, z)
  }  
  return(result)
}
mergesort(c(4,3,2,1,8,7,6,5))
mergesort(c(4,3,2,1,8,7,6,5),TRUE)

```

##section 4.4.1

###exercises 1(a)

```{r}
f=function(x,y)x+y
f_x=function(x,y)1
f_y=function(x,y)1
g=function(x,y)x^2+2*y^2-2
g_x=function(x,y)2*x
g_y=function(x,y)4*y
zero=function(x,y,eps=1e-6){
  while(abs(f(x,y))>=eps || abs(g(x,y))>=eps){
  d=f_x(x,y)*g_y(x,y)-f_y(x,y)*g_x(x,y)
  a=x-(g_y(x,y)*f(x,y)-f_y(x,y)*g(x,y))/d
  b=y-(f_x(x,y)*g(x,y)-g_x(x,y)*f(x,y))/d
  x=a
  y=b
  }
  return(c(x,y))
}
```

###exercises 1(b)

```{r}
zero(-1,1,1e-6)
```

##Chapter 4 

###exercises 1

```{r}
directpoly=function(x,v){
  s=0
  l=length(v)
  for(i in 1:l){
    s=s+x^(l-i)*v[i]
  }
  return (s)
}
directpoly(2,c(1,1,1))
```

###exercises 2

```{r}
hornerpoly=function(x,v){
  l=length(v)
  a=v[l]
  if(l==1) {
    return(a)
  }else{
    for(i in (l-1):1){
      a=a*x+v[i]
    }
  }
  return(a)
}
hornerpoly(2,c(1,1,1))
```

###exercises 3(a)

```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000),c(1, -2, 2, 3, 4, 6, 7)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000),c(1, -2, 2, 3, 4, 6, 7)))
```

###exercises 3(b)

```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000),c(2,17,-3)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000),c(2,17,-3)))
```



